---
title: "WGCNA: Tutorial 1-1"
author: "Sangwon Han"
date: "2019-2-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Data input, cleaning and pre-processing

##1-1. Loadind expression data

Set working directory ad load "LiverFemale3600.csv" file

```{r}
setwd("C:/Users/Han S/Desktop/Han S/WGCNA/tutorials")
suppressMessages(library(WGCNA))
options(stringsAsFactors = FALSE)
femData = read.csv("LiverFemale3600.csv")
dim(femData)
```

```{r}
names(femData)
```

also, we can see data using fix function. (rows = genes, columns = sample or auxiliary information)

remove auxiliary data and transpose the expression data (so, rows = samples, columns = genes)
```{r}
datExpr0 = as.data.frame(t(femData[, -c(1:8)]))
names(datExpr0) = femData$substanceBXH
rownames(datExpr0) = names(femData)[-c(1:8)]
```

##1-2.Missing values and outliers

remove samples with too many missing values. If allOk = FALSE, we should remove the  offending genes
```{r}
gsg = goodSamplesGenes(datExpr0, verbose = 3)
summary(gsg)
```
```{r}
if (!gsg$allOK)
        # print the gene and sample names that were removed:
        if (sum(!gsg$goodGenes)>0)
                printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")))
        if (sum(!gsg$goodSamples)>0)
                printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")))
        
        # remove the offending genes and samples
        datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
```

cluster to see outliers
```{r}
sampleTree = hclust(dist(datExpr0), method = "average")

# open graphic output window 12 (width) * 9 (height) inches 
sizeGrWindow(12, 9)     
par(cex = 0.6)
par(mar = c(0, 4, 2, 0))

# Plot the sample tree
plot(sampleTree, main = "Sample clustering to detect outliers", 
     sub = "", xlab = "", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)

# Plot a cutting line
abline(h = 15, col = "red")
```

Determine cluster under the line
```{r}
clust = cutreeStatic(sampleTree, cutHeight = 15, minSize = 10)
table(clust)
```

clust 1 contains the samples we want to keep
```{r}
keepSamples = (clust == 1)
datExpr = datExpr0[keepSamples, ]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
```


##1-3. Loading clinical trait data
```{r}
traitData = read.csv("ClinicalTraits.csv")
dim(traitData)
```

```{r}
names(traitData)
```

remove columns that hold information we do not need
```{r}
allTraits = traitData[,-c(31,16)]
allTraits = allTraits[, c(2, 11:36)]
dim(allTraits)
```

```{r}
names(allTraits)
```

Form a data frame analogous to expression data that will hold the clinical traits.
```{r}
femaleSamples = rownames(datExpr)
traitRows = match(femaleSamples, allTraits$Mice)        # index of allTraits$Mice which matches with expression data
datTraits = allTraits[traitRows, -1]                    # Extract matched trait data except 1st column(Mice)
rownames(datTraits) = allTraits[traitRows, 1]           # give matched mice name 

collectGarbage()
```

Now, expression data = datExpr, clinical trait data = datTraits
```{r}
# Re-cluster samples
sampleTree2 = hclust(dist(datExpr), method = "average")

# Convert traits to a color representation(white = low, red = high, grey = missing entry)
traitColors = numbers2colors(datTraits, signed = FALSE)

# Plot the sample dendrogram and he colors underneath
plotDendroAndColors(sampleTree2, traitColors, 
                    groupLabels = names(datTraits), main = "Sample dendrogram and trait heatmap")
```

save expression and trait data for use in next steps
```{r}
save(datExpr, datTraits, file = "FemaleLiver-01-dataInput.RData")
```

