---
title: "WGCNA.Tutorial 1-2b"
author: "Sangwon Han"
date: "2019-2-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Set up the R session

```{r}
suppressMessages(library(WGCNA))
options(stringsAsFactors = FALSE)
enableWGCNAThreads()    # allow multi-threading within WGCNA
lnames = load(file = "FemaleLiver-01-dataInput.RData")
lnames
```

#2-2. Step-by-step network construction and module detection
##2-2-1. Choosing the soft-thresholding power

Use the function **pickSoftThreshold**
```{r}
# make a set of powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))

# call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)

sizeGrWindow(9, 5)
par(mfrow = c(1, 2))
cex1 = 0.9

# Scale-free topology fit index as a function of the soft-thresholding power (type = "n" mean no coordinate)
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
        xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit, signed R^2", type = "n",
        main = paste("Scale independence"))                     
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
        labels = powers, cex = cex1, col = "red")
abline(h = 0.90, col = "red")


# Plot mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab = "Soft Threshold (power)", ylab = "Mean connectivity", type = "n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels = powers, cex = cex1, col = "red")

```

We choos the power 6.


##2-2-2. Co-expression similarity and adjacency

Calculate adjacency (softpower = 6)
```{r}
softPower = 6
adjacency = adjacency(datExpr, power = softPower)
```


##2-2-3. Topological Overlap Matrix (TOM)

Transform the adjacency into TOM and calculate the dissimilarity: to minimize effect of noise
```{r}
TOM = TOMsimilarity(adjacency)
dissTOM = 1-TOM
```


##2-2-4. Clustering using TOM
```{r}
geneTree = hclust(as.dist(dissTOM), method = "average")
sizeGrWindow(12,9)
plot(geneTree, xlab = "", sub="", main = "Gene clustering on TOM-based dissimilarity", label = FALSE, hang = 0.04)
```

For branch cutting, use the package dynamicTreeCut
```{r}
minModuleSize = 30      # We like large module

# Module identification using dynamic cut tree
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                            deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize)
table(dynamicMods)
```

It returns 22 modules (1: largest to 22: smallest. 0 = unassigned genes)
Convert numeric labels into colors.

```{r}
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
```

Plot the dendrogram and color underneath
```{r}
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
```



##2-2-4. Merging modules with similar expression profiles

Cluster module eigengenes. Merge using a height cut of 0.25 (correlation = 0.75).
```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes

# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs)

# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average")
sizeGrWindow(7,6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
MEDissThres = 0.25
abline(h = MEDissThres, col = "red")

# Merge using a height cut of 0.25 
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
mergedColors = merge$colors
mergedMEs = merge$newMEs
```


Plot the gene dendrogram agian with original and merged module colors underneath
```{r}
sizeGrWindow(12, 9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
```


Now, save the  relevant variables for subsequent analysis.
```{r}
# Rename to moduleColors
moduleColors = mergedColors 

# Construct numeric labels corresponding to the colors
colorOrder = c("grey", standardColors((50)))
moduleLabels = match(moduleColors, colorOrder)-1
MEs = mergedMEs

# Save module colors and labels
save(MEs, moduleLabels, moduleColors, geneTree, file = "FemaleLiver-02-networkConstruction-stepByStep.RData")
```



