---
title: "WGCNA: Tutorial 1-2a"
author: "Sangwon Han"
date: "2019-2-27"
output: html_document
---

#1 Set up the R session

```{r}
suppressMessages(library(WGCNA))
options(stringsAsFactors = FALSE)
enableWGCNAThreads()    # allow multi-threading within WGCNA
lnames = load(file = "FemaleLiver-01-dataInput.RData")
lnames
```

#2-1. Automatic network construction and module detection
##2-1-1. Choosing the soft-thresholding power

Use the function **pickSoftThreshold**
```{r}
powers = c(c(1:10), seq(from = 12, to = 20, by =2))     # make a set of powers

# call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
str(sft)
```

Plot the results                        
```{r}
sizeGrWindow(9, 5)
par(mfrow = c(1, 2))
cex1 = 0.9

# Scale-free topology fit index as a function of the soft-thresholding power (type = "n" mean no coordinate)
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
        xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit, signed R^2", type = "n",
        main = paste("Scale independence"))                     
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
        labels = powers, cex = cex1, col = "red")
abline(h = 0.90, col = "red")


# Plot mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab = "Soft Threshold (power)", ylab = "Mean connectivity", type = "n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels = powers, cex = cex1, col = "red")
```

We choose power 6, for which scale-free topology fit index curve flatten with high value (0.90)



##2-1-2. One-step network construction and module detection

we set relatively large minimum module size (**minModuleSize = 30**), medium sensitivity (**deepSplit = 2**).
**maxBlockSize**  = 5000 in deflaut, so dataset with > 5000 probes will lead to errors.
(4GB memory can handle 10,000 probes, 16GB -> 20,000, 32GB -> 30000)

- **mergeCutHeight** = the threshold for merging of modules
- **numericLables** = return numeric (not color) labels for modules 
- **saveTOMs** = save Topological Overlap Matrix
```{r}
net = blockwiseModules(datExpr, power = 6,
                       TOMtype = "unsigned", minModuleSize = 30,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "femaleMouseTOM",
                       verbose = 3)
```

- net$colors = module assignment
- net$MEs = module eigengenes
```{r}
str(net)
```

There are 18 modules in descending size order (1~18, 0 = outside of all modules)
```{r}
table(net$colors)
```

The hierarchical clustering dendrogram used for the module identification
```{r}
# open a graphics window
sizeGrWindow(12, 9)

# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)

# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                "Module colors",
                dendroLabels = FALSE, hang = 0.03,
                addGuide =  TRUE, guideHang = 0.05)
```
**recutBlockwiseTrees**: to change the treecut, module membership, module merging criteria

Now, save the module assignment and module eigengene information for subsequent analysis.
```{r}
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs
geneTree = net$dendrograms[[1]]
save(MEs, moduleLabels, moduleColors, geneTree, file = "FemaleLiver-02-networkConstruction-auto.RData")
```

